Detection of Oscillations in Aldebaran with Ground-Based Observations

### 2017-09-26 paper/aldebaran.tex
small changes

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. Following the proposal of \citet{struverv} several groups began to search for planets by the radial velocity (RV) method, looking for Doppler shifts from the star's reflex motion around the common centre of mass, culminating in the discovery of 51~Peg~b \citep{51peg}, the first exoplanet to be recognized as such. Before this, \citet{hatzes1993} noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran. After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

@@

In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods.

Aldebaran was therefore observed with the \ktwo Mission \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.

@@

We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicitiies $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.  %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.  
Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models, we compute the luminosity and radius evolution of the planet along the main sequence. $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr. 
We note that this implies that the planet at a semi-major axis of $1.507 \pm 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86 \pm 0.27$ to $1.55 \pm 1.0$ times that of the Earth today. Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments for some of their previous evolution, now long-since destroyed by their host star's evolution away from the main sequence.

>>>>

Aldebaran, or $\alpha$~Tauri, is a well-known first-magnitude naked-eye red giant star, and has long been the subject of astronomical investigations. By looking for Doppler shifts from the star's reflex motion around the common centre of mass with a companion planet \citep[the radial velocity or RV method;][]{struverv}, . While the hot Jupiter 51~Peg~b \citep{51peg} was the first exoplanet to be recognized as such, before this, \citet{hatzes1993} had noted RV variations in Pollux \citep[$\beta$~Gem; subsequently confirmed as a planet:][]{betgemconf,betgemconf2}, Arcturus ($\alpha$~Boo, unconfirmed), and Aldebaran. After further investigation by \citet{Hatzes1998}, \citet{Hatzes2015} now claim a firm RV detection of a planetary-mass companion Aldebaran~b, with a period of $628.96 \pm 0.90$~d.

@@

In order to verify the results of the novel analysis presented above, we sought to obtain an independent detection of the oscillations of Aldebaran and compare the frequencies determined with the two methods. 
Aldebaran was therefore observed with \ktwo \citep{howell14}, a two-wheeled revival of the \kepler Mission \citep{2010sci...327..977b}, under Guest Observer Program 130471 in Campaign~13, from 2017-03-08 to 2017-05-27.

@@

We note that with the determination of the stellar mass it is possible to infer the parameters Aldebaran had while it was on the main sequence. We conduct a Monte Carlo simulation, drawing masses randomly from a Gaussian distribution $M \sim \mathcal{N}(1.16,0.07)$ and metallicitiies $[\text{Fe}/\text{H}] \sim \mathcal{N}(-0.15,0.2)$ \citep{decin2003}  to predict the luminosity of the main-sequence progenitor, and the semi-major axis of the planet's orbit.  %Using the scaling relations of \citep{eker2015}, we find that Aldebaran would have had a luminosity $L_* = 2.0 \pm 0.37~\lsun$ on the main sequence.  
Using the Mesa Isochrones and Stellar Tracks \citep[MIST:][]{mist0,mist1} models via the \textsc{isochrones} package \citep{isochrones}, we compute the luminosity and radius evolution of the planet along the main sequence. $L$ evolves from $2.0 \pm 0.7$~\lsun at 0.5~Gyr to $3.5 \pm 2.3$~\lsun at 4.5~Gyr. 
We note that this implies that the planet at a semi-major axis of $1.507 \pm 0.03$~AU would have been subject to an insolation comparable to that of the Earth, evolving from 0.5~Gyr to 4.5~Gyr from $0.86 \pm 0.27$ to $1.55 \pm 1.0$ times that of the Earth today. Subject as well as this to the great uncertainties of the planet's orbital evolution and albedo, this planet and any of its moons may well have had temperate environments for some of their previous evolution, now long-since destroyed by their host star's evolution away from the main sequence.
